The finite representation property for some reducts of relation algebras

### 2021-06-15 filtration-stuff/NewText.tex
more facts about definable filtrations for fusions

\author{Daniel Rogozin}
\date{}
\title{Notes on filtration of logics containing {\bf K5}}
\begin{document}

@@

@@

@@

\begin{lemma}\label{extension}
  Let $\Gamma$ be a finite and $\operatorname{Sub}$-closed set of formulas and $\mathcal{M} = \langle W, R, \vartheta \rangle$. Consider $\Gamma' = \Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$. Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$. Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{E}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
\end{lemma}

@@

  \xymatrix{
  &&& \hat{z} & \ar[l]_{R_{\Delta}^{min}} \ar[r]^{R'} \widehat{u_1} & \ar[r]^{R'} \widehat{u_2} \ar[r]^{R'} & \dots \ar[r]^{R'} & \widehat{u_{n - 1}} \ar[r]^{R'} & \widehat{u_{n}} \ar[r]^{R_{\Delta}^{min}} & \hat{x}
  }
  where $R'$ is either $R_{\Delta}^{min}$ or its converse. We have $\mathcal{M}, z \models \Box \psi$, since $\mathcal{M}, x \models \Box \psi$ implies $\mathcal{M}, u_{n} \models \Diamond \Box \psi$. After that we apply the following property of ${\bf K}5$-models:
  \begin{center}
    For each $a, b \in M_i$ such that $a R_i b$ we have $\mathcal{M}_i, a \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}_i, b \models \Diamond \Box p_{\psi}$
  \end{center}
  Note that we always stay within $\Delta$ since $\Gamma' \subseteq \Delta$.
  Finally, we conclude $\mathcal{M}, x \models \psi$ from IH and $\mathcal{M}, z \models \Box \psi$.
\end{proof}

@@

  \begin{enumerate}
  \item ${\bf K}5 * {\bf K}5$ admits definable filtrations.
  \item ${\bf K}5 * \dots * {\bf K}5$ admits definable filtrations.
  \item If $\mathcal{L}_1, \dots, \mathcal{L}_n$ admit strict filtrations, then ${\bf K}5 * \dots * \dots * {\bf K}5 * \mathcal{L}_1 * \dots * \mathcal{L}_n$
  \end{enumerate}
\end{theorem}

@@

  \item Likewise
  \end{enumerate}
\end{proof}

\begin{theorem}
  Let $\mathcal{L}_1$ and $\mathcal{L}_2$ be modal logics such that $\operatorname{Fr}(\mathcal{L}_1)$ and $\operatorname{Fr}(\mathcal{L}_2)$ admit filtrations, so does $\operatorname{Fr}(\mathcal{L}_1 * \mathcal{L}_2$).
\end{theorem}

\begin{proof}

>>>>

\author{Daniel Rogozin}
\date{}
\title{Notes on filtrations for logics that contain {\bf K5}}
\begin{document}

@@

\section{Preliminaries}

@@

\section{Filtrations for {\bf K5}}

@@

\begin{lemma}\label{extension}
  $ $

  \begin{itemize}
  \item Let $\Gamma$ be a finite and $\operatorname{Sub}$-closed set of formulas and $\mathcal{M} = \langle W, R, \vartheta \rangle$. Consider $\Gamma' = \Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$. Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$. Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{E}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
  \item The same for ${\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$
  \end{itemize}
\end{lemma}

@@

  \xymatrix{
  &&& \hat{z} & \ar[l]_{R_{\Delta}^{min}} \ar[r]^{R'} \widehat{u_1} & \ar[r]^{R'} \widehat{u_2} \ar[r]^{R'} & \dots \ar[r]^{R'} & \widehat{u_{n - 1}} \ar[r]^{R'} & \widehat{u_{n}} \ar[r]^{R_{\Delta}^{min}} & \hat{x}
  }
  where $R'$ is either $R_{\Delta}^{min}$ or its converse. We have $\mathcal{M}, z \models \Box \psi$, since $\mathcal{M}, x \models \Box \psi$ implies $\mathcal{M}, u_{n} \models \Diamond \Box \psi$. After that we apply the following property of ${\bf K}5$-models:
  \begin{center}
    For each $a, b \in M_i$ such that $a R_i b$ we have $\mathcal{M}_i, a \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}_i, b \models \Diamond \Box p_{\psi}$
  \end{center}
  Note that we always stay within $\Delta$ since $\Gamma' \subseteq \Delta$.
  Finally, we conclude $\mathcal{M}, x \models \psi$ from IH and $\mathcal{M}, z \models \Box \psi$.

  The second item is shown similarly.
\end{proof}

@@

  \begin{enumerate}
  \item ${\bf K}5 * {\bf K}5$ admits definable filtrations.
  \item ${\bf K}5 * \dots * {\bf K}5$ admits definable filtrations.
  \item If $\mathcal{L}$ admits strict filtrations, then ${\bf K}5 * \mathcal{L}$ admits definable filtrations
  \item If $\mathcal{L}_1, \dots, \mathcal{L}_n$ admit strict filtrations, then ${\bf K}5 * \dots * \dots * {\bf K}5 * \mathcal{L}_1 * \dots * \mathcal{L}_n$
  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$  (here and below), then $\mathcal{L} * \mathcal{L}$ admits definable filtrations.
  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$ and $\mathcal{L}_1$ a logic that admits strict filtrations, then $\mathcal{L} * \mathcal{L}_1$
  \end{enumerate}
\end{theorem}

@@

  \item Likewise
  \item The argument is the same as in the proof of the first item, except for {\bf Claim} 1 that has the following formulation:
  Let $\widehat{R}_1 = (R^{min}_{\Delta})^{E}$ and $\widehat{R}_2 = (R^{min}_{\Delta})^{\mathcal{L}_1}$
  Define a valuation as usual as $\widehat{\vartheta(p)} = \{ [x]_{\sim_i} \: | \: \mathcal{M}_i, x \models p \}$ for $p \in \Delta_1$, define $\widehat{\mathcal{M}}_1 = \langle \widehat{W}, \widehat{R}_1, \widehat{\vartheta} \rangle$ and $\widehat{\mathcal{M}}_2 = \langle \widehat{W}, \widehat{R}_2, \widehat{\vartheta} \rangle$.

  Then $\widehat{\mathcal{M}}_1 \models {\bf K}5$ and $\widehat{\mathcal{M}}_1 \models {\bf L}$ and $\widehat{\mathcal{M}}_i$ is a filtration of $\mathcal{M}_i$ through $\Delta$.

  \item Likewise
  \item The argument is similar to the proof of first item, but filtrations are slightly different.
  Let $\mathcal{M} = \langle W, R_1, R_2, \vartheta \rangle$ be a $\mathcal{L} * \mathcal{L}$ model and $\Gamma$ a $\operatorname{Sub}$-closed set of formulas. As above, we define a set $V$ and a model $\mathcal{M}'$. Define extensions of $\Gamma$ and $V$:

  \begin{center}
    $\Gamma' = \Gamma \cup \{ \Diamond_1 \Diamond_1 \psi \: | \: \Diamond_1 \psi \in \Gamma \} \cup \{ \Diamond_2 \Diamond_2 \psi \: | \: \Diamond_2 \psi \in \Gamma \}$

    $\Delta = V \cup \operatorname{Sub}(\{ \Diamond \Diamond p_{\psi} \: | \: \Diamond \psi \in \Gamma, i = 1,2 \})$
  \end{center}
  As above $\sim_{\Delta} = \Gamma'$ and $\widehat{\mathcal{M}'} = \langle W /\sim_{\Delta}, \widehat{R_i}, \widehat{\vartheta} \rangle$ are filtrations of reducts of $\mathcal{M}'$ through $\Delta$. Then $\widehat{\mathcal{M}} = \langle W /\sim_{\Delta}, \widehat{R_1}, \widehat{R_2}, \widehat{\vartheta} \rangle$ is a required filtration of the original $\mathcal{M}$.
  \item Extend $\Gamma$ with $\{ \Diamond_i \Diamond_i \psi \: | \: \Diamond_i \psi \in \Gamma, i = 1,2 \}$ and
  $V$ with $\{\Diamond \Diamond p_{\psi} \: | \: \Diamond_i {\psi}, i = 1, 2\}$
  \end{enumerate}
\end{proof}

\begin{theorem}
  Let $\mathcal{L}_1$ and $\mathcal{L}_2$ be modal logics that admit definable filtrations. If $\operatorname{Mod}(\mathcal{L}_1)$ and $\operatorname{Mod}(\mathcal{L}_2)$ admit definable filtrations, so does $\operatorname{Mod}(\mathcal{L}_1 * \mathcal{L}_2$).
\end{theorem}

\begin{proof}
  Let $\mathcal{M} = \langle W, R_1, R_2, \vartheta \rangle$ be an $\mathcal{L}_1 * \mathcal{L}_2$.
  We define a notation $\nabla = \{ \neg \Diamond, \Diamond \neg, \Diamond \}$.

  Both logics admit definable filtrations, so for every finite $\operatorname{Sub}$-closed set $\Gamma$ and for every $\mathfrak{M}$, an $\mathcal{L}_1$-model (or an $\mathcal{L}_2$ one) there exists there exists $\Delta$, a extension of $\Gamma$ having the form:
  \begin{center}
    $\Delta_1 = \Gamma \cup \operatorname{Sub}\{ \nabla_1 \nabla_2 \dots \nabla_n \Diamond \psi \: | \: \Diamond \psi \in \Gamma \}$ (for $\mathcal{L}_1$)

    $\Delta_2 = \Gamma \cup \operatorname{Sub}\{ \nabla_1 \nabla_2 \dots \nabla_k \Diamond \psi \: | \: \Diamond \psi \in \Gamma \}$ (for $\mathcal{L}_2$)
  \end{center}
  such that $\widehat{\mathfrak{M}} = \langle W / \sim_{\Delta_i}, \widehat{R}, \vartheta \rangle$ is a filtration of $\mathfrak{M}$ through the corresponding $\Delta_i$.

  Let $V$ be a set of fresh variables indexed over $\Gamma$ as in the proof for a fusion of ${\bf K5}$ with something else. Let $\mathcal{M}'$ be a model defined as previously. We extend $V$ and $\Gamma$ in the following way:

\begin{center}
  $\Gamma' = \Gamma \cup \operatorname{Sub}\{ {\nabla_1}_1 {\nabla_2}_1 \dots {\nabla_n}_1 \Diamond_1 \psi \: | \: \Diamond_1 \psi \in \Gamma \} \cup \operatorname{Sub}\{ {\nabla_1}_2 {\nabla_2}_2 \dots {\nabla_n}_2 \Diamond_2 \psi \: | \: \Diamond_2 \psi \in \Gamma \}$

  $\Delta = V \cup \operatorname{Sub}\{ \nabla_1 \nabla_2 \dots \nabla_n \Diamond p_{\psi} \: | \: {\nabla_{n + 1}}_1 \psi \in \Gamma' \} \cup \operatorname{Sub}\{ \nabla_1 \nabla_2 \dots \nabla_k \Diamond p_{\psi} \: | \: \Diamond_2 \psi \in \Gamma \}$.
\end{center}
By the construction, $\sim_{\Gamma'} = \sim_{\Delta}$. So we have filtrations for the corresponding reducts of $\mathcal{M}'$ through $\Delta$ as well as for the original $\mathcal{M}$.
\end{proof}

### 2021-06-15 frp-paper/Paper.tex
more facts about definable filtrations for fusions

Let $\mathcal{A}$ be a residuated semigroup and $\mathcal{Q}_{\mathcal{A}}$ is a quantale of Galois closed subsets of $\mathcal{A}$. $\widehat{\mathcal{Q}_{\mathcal{A}}}$ is the corresponding relational quantale. Let us define an interpretation $R : \mathcal{A} \to \widehat{\mathcal{Q}_{\mathcal{A}}}$ such that:

@@

  \begin{enumerate}
    \item If $\mathcal{A}$ is representable then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
    \item If $|\mathcal{A}| = \omega$ and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
  \end{enumerate}
\end{lemma}

>>>>

Let $\mathcal{A}$ be a residuated semigroup and $\mathcal{Q}_{\mathcal{A}}$ a quantale of Galois closed subsets of $\mathcal{A}$. $\widehat{\mathcal{Q}_{\mathcal{A}}}$ is the corresponding relational quantale. Let us define an interpretation $R : \mathcal{A} \to \widehat{\mathcal{Q}_{\mathcal{A}}}$ such that:

@@

  \begin{enumerate}
    \item If $\mathcal{A}$ is representable then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
    \item If $|\mathcal{A}| \leq \omega$ and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
  \end{enumerate}
\end{lemma}
